---
title: "Homework 1"
output:
  pdf_document: default
  html_notebook: default
date: "9/16/2020"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy=TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(pROC)
```


### Question 1
#### Use the  airbnb_data.csv provided and answer the following questions (Q1a. b. c. and d.) on Linear Regression.

Importing the data:
```{r}
airbnb <- read.csv("airbnb_data.csv", header = TRUE)
summary(airbnb)
```

#### a) Fit a multiple linear regression model using price as the response variable and all others as predictor variables (Note: remove 'id' columns). Which variables are statistically significant in determining the price?

```{r}
# Removing the id columns
model_data = select(airbnb, -c('room_id', 'survey_id', 'host_id', 'city'))

# Fitting & outputting the model
airbnb_model <- lm(price ~ room_type + reviews + overall_satisfaction + accommodates + bedrooms, data = model_data)
summary(airbnb_model)
```
Based on the p value, the room type (room_typePrivate room and room_typeShared room columns in the data set) and reviews are not statistically significant as their p value is greater than 0.05. The remaining predictors (overall_satisfaction, accomodates and bedrooms) are the statisically significant predictors of price.

##### b) Interpret the coefficients for predictors: room type(Shared Room), bedrooms?
Room Type (Shared Room): if the Airbnb property is a shared room, the price will decrease by $76.68 
Bedrooms: the addition of one extra bedroom will increase the price by $85.65

##### c) Predict the price (nearest dollar) for a listing with the following factors: bedrooms = 1, accommodates = 2, reviews = 70, overall_satisfaction = 4, and room_type= 'Private room'.
```{r}
listing = data.frame(bedrooms = 1, accommodates = 2, reviews = 870, overall_satisfaction = 4, room_type = "Private room" )
predict(airbnb_model, newdata = listing)
```
Based on the specs of the given listing and the model fitted in part a, the price of this listing is $75.

##### d) Identify outliers using Cook's distance approach. Remove points having Cook's distance > 1. Rerun the model after the removal of these points and print the summary.

```{r message = FALSE}
# Calculate Cook's distance
cooks_dist <- cooks.distance(airbnb_model)

# Graphing Cook's distance
sample_size <- nrow(model_data)
plot(cooks_dist, pch="*", cex=2, main="Influential Data Points by Cook's Distance")
abline(h = 1, col="red")  # add cutoff line

# Identifying whic points have Cook's distance > 1 and removing them from the data set
influential <- as.numeric(names(cooks_dist)[(cooks_dist > 1)])
model_data2 <- model_data[-influential,]

# Creating a new model based on the data without the influential points 
airbnb_model2 <- lm(price ~ room_type + reviews + overall_satisfaction + accommodates + bedrooms, data = model_data2)
summary(airbnb_model2)
```
After removing the influential points, the room type is now statistically significant. The R squared and adj. R squared value of this model is better than the previous model that includes the influential points.


### Question 2
#### Use the direct_marketing.csv provided and answer the following questions (Q2.a and Q2.b) on Linear Regression.

Importing the data:
```{r message = FALSE}
dir_market <- read.csv("direct_marketing.csv", header = TRUE)
summary(dir_market)
```
#### Create indicator variables for the ‘History’ column. Considering the base case as None (i.e., create Low, Medium and High variables with 1 denoting the positive case and 0 the negative) and few additional variables LowSalary, MediumSalary and HighSalary based on the customer history type i.e., MediumSalary = Medium*Salary etc.

```{r, tidy=TRUE}
# Creating indicator variables for History column
dir_market<- dir_market %>% mutate(history_low = ifelse(History=="Low",1,0)) %>% mutate(history_med = ifelse(History=="Medium",1,0))%>% mutate(history_high = ifelse(History=="High",1,0))

# Creating interaction terms
dir_market<- dir_market %>% mutate(LowSalary = history_low*Salary) %>% mutate(MediumSalary=history_med*Salary) %>% mutate(HighSalary = history_high*Salary)

summary(dir_market)
```

#### a) Fit a multiple linear regression model using AmountSpent as the response variable and the indicator variables along with their salary variables as the predictors.

```{r}
# Creating the model
dm_model <- lm(AmountSpent ~ history_low + history_med + history_high + LowSalary + MediumSalary + HighSalary, data = dir_market)

summary(dm_model)
```

#### b) What is the amount spent by a customer for each historic type provided their salary is $10,000 based on the model constructed in part a?

```{r}
none_sal = data.frame(history_low = 0, history_med = 0, history_high = 0, LowSalary = 0, MediumSalary = 0, HighSalary= 0)
low_sal = data.frame(history_low = 1, history_med = 0, history_high = 0, LowSalary = 10000, MediumSalary = 0, HighSalary= 0)
med_sal = data.frame(history_low = 0, history_med = 1, history_high = 0, LowSalary = 0, MediumSalary = 10000, HighSalary= 0)
high_sal = data.frame(history_low = 0, history_med = 0, history_high = 1, LowSalary = 0, MediumSalary = 0, HighSalary=10000)

none_amountspent = predict(dm_model, newdata = none_sal)
low_amountspent = predict(dm_model, newdata = low_sal)
med_amountspent = predict(dm_model, newdata = med_sal)
high_amountspent = predict(dm_model, newdata = high_sal)

cat(none_amountspent,'\n',low_amountspent,'\n',med_amountspent,'\n',high_amountspent)
```
Given a salary of $10,000, a customer with no history will spend $124, a customer with low history will spend $30, a customer with medium history will spend $84 and a customer with high history will spend $92.

#### Use the airbnb_data.csv Preview the document provided and answer the following questions (Q2.c and Q2.d) on Linear Regression. Perform Log transformation for the variables price and overall_satisfaction, make necessary transformations suggested in the class.

#### c) Fit all four models i.e., linear-linear, linear-log, log-linear, and log-log regression models using price as the response variable and overall_satisfaction as the predictor.

```{r}
# Generating the four lin/log models. Since overall_statisfaction has values of 0, I use the function log1p instead of log to add 1 to every data point
lin_lin <- lm(price ~ overall_satisfaction, data = airbnb)
lin_log <- lm(price ~ log1p(overall_satisfaction), data = airbnb)
log_lin <- lm(log(price) ~ overall_satisfaction, data = airbnb)
log_log <- lm(log(price) ~ log1p(overall_satisfaction), data = airbnb)
```
#### d) Which of the four models has the best R2? Do you have any comments on the choice of the dependent variable?

```{r}
cat("R Squared Values:",'\n',"Linear-Linear:",summary(lin_lin)$r.squared,'\n',"Linear-Log:",summary(lin_log)$r.squared,'\n',"Log-Linear",summary(log_lin)$r.squared,'\n',"Log-Log", summary(log_log)$r.squared,'\n')

summary(lin_log)
```
The model with the best R squared value is the linear-log model. However, the R squared value for all 4 models is very low. Even though overall satisfaction appears to be statistically significant, the low R squared value is an indication that overall satisfaction does not explain much of the variation in the price and another dependant variable should replace overall satisfaction as a predictor or be addded to the model.

### Question 3
#### The attached titanic_data.csv has been cleaned to remove all rows which contain missing values. We will perform logistic regression on this cleaned dataset.
###### The dataset contains the following columns:
###### ‘Name’ - Passenger Name - factor
###### ‘PClass’ - Passenger Class (1st, 2nd, 3rd) - factor
###### ‘Age’ - Passenger Age - number
###### ‘Sex’ - Passenger Sex – female, male
###### ‘Survived’ – 1 if passenger survived, 0 if not - number
#### After converting the survived variable to be a factor with two levels, 0 and 1, perform logistic regression on the dataset using ‘Survived’ as the response and ‘Sex’ as the explanatory variable.

#### a) Display the model summary.

```{r}
# Importing the data
titanic <- read.csv("titanic_data.csv", header = TRUE)

# Converting Survived to a factor and Sex to an indicator variable
titanic$Survived <- as.factor(titanic$Survived)
titanic<- titanic %>%
  mutate(Sex = ifelse(Sex=="female",1,0))

# Fitting the model
titanic_model <- glm(Survived ~ Sex, data = titanic, family = 'binomial')
summary(titanic_model)
```
#### b) What does the value of the intercept coefficient represent in this model?

The intercept coefficient of -1.3545 is the log odds for the base case of Sex = 0 or Sex = male. 
It is used to calculate the probability that Survived = 1, given that a person is male in p(x) = exp(b0+b1x)/[1 + exp(b0+b1x)]

#### c) Determine the probability of survival for females.
```{r}
b0 = coef(titanic_model)[1]
b1 = coef(titanic_model)[2]
x = 1
p_female = exp(b0+b1*x)/(1 + exp(b0+b1*x))

print(p_female)
```
The probability of survival for females is 75.35%.

#### d) Determine the probability of survival for males.
```{r}
b0 = coef(titanic_model)[1]
b1 = coef(titanic_model)[2]
x = 0
p_male = exp(b0+b1*x)/(1 + exp(b0+b1*x))

print(p_male)
```
The probability of survival for males is 20.51%.